### 2009-04-03 talk/icooolps2009-dotnet/benchmarks.tex
briefly describe TLC

\anto{XXX: we need to give an overview of TLC}

In section \ref{sec:tlc-properties}, we saw that TLC provides most of the
features that usually make dynamically typed language so slow, such as
\emph{stack-based interpreter}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

@@

\begin{enumerate}
\item By plain interpretation, without any jitting.
\item With the JIT enabled: this run includes the time spent by doing the
  compilation itself, plus the time spent by running the produced code.
\item Again with the JIT enabled, but this time the compilation has already
  been done, so we are actually measuring how good is the code we produced.
\end{enumerate}

>>>>

To measure the performances of the CLI JIT backend, we wrote a simple virtual
machine for a dynamic toy languaged, called \emph{TLC}.

The design goal of the language is to be very simple (the interpreter of the
full language consists of about 600 lines of RPython code) but to still have
the typical properties of dynamic languages that make them hard to
compile. TLC is implemented with a small interpreter that interprets a custom
bytecode instruction set. Since our main interest is in the runtime
performance of the interpreter, we did not implement the parser nor the
bytecode compiler, but only the interpreter itself.

Despite being very simple and minimalistic, \lstinline{TLC} is a good
candidate as a language to test our JIT generator, as it has some of the
properties that makes most of current dynamic languages (e.g. Python) so slow:

\begin{itemize}

\item \textbf{Stack based interpreter}: this kind of interpreter requires all the operands to be
  on top of the evaluation stack.  As a consequence programs spend a lot of
  time pushing and popping values to/from the stack, or doing other stack
  related operations.  However, thanks to its simplicity this is still the
  most common and preferred way to implement interpreters.

\item \textbf{Boxed integers}: integer objects are internally represented as
  an instance of the \lstinline{IntObj} class, whose field \lstinline{value}
  contains the real value.  By having boxed integers, common arithmetic
  operations are made very slow, because each time we want to load/store their
  value we need to go through an extra level of indirection.  Moreover, in
  case of a complex expression, it is necessary to create many temporary
  objects to hold intermediate results.

\item \textbf{Dynamic lookup}: attributes and methods are looked up at
  runtime, because there is no way to know in advance if and where an object
  have that particular attribute or method.
\end{itemize}

@@

\begin{enumerate}
\item By plain interpretation, without any jitting (\emph{Interp}).
\item With the JIT enabled: this run includes the time spent by doing the
  compilation itself, plus the time spent by running the produced code (\emph{JIT}).
\item Again with the JIT enabled, but this time the compilation has already
  been done, so we are actually measuring how good is the code we produced (\emph{JIT 2}).
\end{enumerate}
